# Most hamful stroms and other weathers in the U.S from 1950 - 2011

#### Jaehyun Shin, 19. Dec. 2014, Reproduciabl Research Course Projec
======

## Synopsis
The aim of this report is to analyze top harmful storms and severe weather events in the U.S between the years 1950 and 2011. 

Storms and other severe weather events can cause both public health and economic problems for communities and municipalities. Many severe events can result in fatalities, injuries, and property damage, and preventing such outcomes to the extent possible is a key concern.

Source of this project is from the U.S. National Oceanic and Atmospheric Administration's (NOAA) storm database. This database tracks characteristics of major storms and weather events in the United States, including when and where they occur, as well as estimates of any fatalities, injuries, properties, crops damage.

## Data Processing
======
### prepare libraries and data  
* [Storm Data](https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2FStormData.csv.bz2)
* [Storm Data Document](https://d396qusza40orc.cloudfront.net/repdata%2Fpeer2_doc%2Fpd01016005curr.pdf)
* [Strom Data FAQ](https://d396qusza40orc.cloudfront.net/repdata%2Fpeer2_doc%2FNCDC%20Storm%20Events-FAQ%20Page.pdf)
```{r preparesource, message=FALSE, cache=TRUE}
library(data.table)
library(dplyr)
library(ggplot2)

# Assum Strom Data in current local path
data = data.table(read.csv(bzfile('repdata-data-StormData.csv.bz2')))
```

### Processing
```{r processing, warning=FALSE, cache=TRUE}

# Make padding data by EXP data
calcPadding = function(EXP){
    ifelse(grepl('^\\d$', EXP), (as.numeric(EXP) * 10),
	ifelse(grepl('^[H|h]$', EXP), 100,
	ifelse(grepl('^[K|k]$', EXP), 1000,
	ifelse(grepl('^[M|m]$', EXP), 1000000,
	ifelse(grepl('^[B|b]$', EXP), 1000000000,
	0)))))
}

pData = select(data, EVTYPE, FATALITIES, INJURIES)
pData = mutate(data, EVTYPE = as.factor(EVTYPE))
eData = select(data, EVTYPE, PROPDMG, PROPDMGEXP, CROPDMG, CROPDMGEXP)
eData = mutate(eData, EVTYPE = as.factor(EVTYPE), PROPDMG =  PROPDMG * calcPadding(PROPDMGEXP), CROPDMG = CROPDMG * calcPadding(CROPDMGEXP))
```

#### Processing about population health
```{r p_processing, warning=FALSE, cache=TRUE}
pResult = group_by(pData, EVTYPE) %>%
	summarise(SUM_FATALITIES = sum(FATALITIES), SUM_INJURIES = sum(INJURIES)) %>%
	arrange(desc(SUM_FATALITIES + SUM_INJURIES)) %>%
	slice(1:10) %>%
	mutate(EVTYPE = factor(EVTYPE, levels = EVTYPE))
```

#### Processing about economic consequences
```{r e_processing, warning=FALSE, cache=TRUE}
eResult = group_by(eData, EVTYPE) %>%
	summarise(SUM_PROPDMG = sum(PROPDMG), SUM_CROPDMG = sum(CROPDMG)) %>%
	arrange(desc(SUM_PROPDMG + SUM_CROPDMG)) %>%
	slice(1:10) %>%
	mutate(EVTYPE = factor(EVTYPE, levels = EVTYPE))
```
  
## Results
======
### Most harmful with respect to population health

#### Table
```{r}
pResult
```

#### Plot
```{r fig.width=13, fig.height=7, dev="svg"}
pMelt = melt(pResult, id.vars= 'EVTYPE')
p1 = ggplot(pMelt, aes(x = EVTYPE, y = value, fill = variable))
p1 = p1 + geom_bar(stat = 'identity')
p1 = p1 + xlab('Event Type') + ylab('Total') + theme(axis.text.x = element_text(angle = 45, hjust = 1))
p1 = p1 + ggtitle('Top 10 Harmful Stroms and Severe Weather Events')
p1 = p1 + theme(plot.title = element_text(vjust=2, size=rel(2)))
p1
```

### Events have the greatest economic consequences

#### Table
```{r}
eResult
```

#### Plot

```{r fig.width=13, fig.height=7, dev="svg"}
eMelt = melt(eResult, id.vars = 'EVTYPE')
p2 = ggplot(eMelt, aes(x = EVTYPE, y = value, fill = variable))
p2 = p2 +  geom_bar(stat = 'identity')
p2 = p2 + xlab('Event Type') + ylab('Damage') + theme(axis.text.x = element_text(angle = 45, hjust = 1))
p2 = p2 + ggtitle('Top 10 Economic Damages By Stroms and Severe Weather Events')
p2 = p2 + theme(plot.title = element_text(vjust=2, size=rel(2)))
p2
```

## Conclusion
======
From my reseach, I conclude that tornado is most harmful to Fatalities and Injuries and flood occure greatest damage to properties and drought occure greatest damage to crops
